---
title: "p8105_hw6_hw2849"
author: "Haotian Wu"
date: "11/30/2021"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(modelr)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%",
	dpi = 300
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## Problem 1

##### Load and Clean Data 

```{r, data import and tidy}
birthweight = read_csv("./birthweight.csv") %>% 
  janitor::clean_names() %>% 
  drop_na() %>%
  mutate( ## recode factor variables
    babysex = recode_factor(babysex, "1" = "male", "2" = "female"),
    frace = recode_factor(frace, "1" = "White", "2" = "Black", "3" = "Asian", "4" = "Puerto Rican", "8" = "Other"),
    mrace = recode_factor(mrace, "1" = "White", "2" = "Black", "3" = "Asian", "4" = "Puerto Rican"),
    malform = recode_factor(malform, "0" = "absent", "1" = "present" )
    ) %>%
  print()
```

We converted variables into correct types. For example, `babysex`, `frace`, `mrace`, and `malform` were present in numeric, and after tidying, these variables are recoded into factor. Missing data are dropped. After tidy, there are 4342 observations with 20 variables included in the data set. 

##### Regression Models

Evidences have shown biological and environmental behaviors of parents are associated with infant birth weight. We hypothesized parents' races, family incomes, pre-pregnancy weight and BMI, mothers' weight change during pregnancy, and smoking behavior would be factors that affect infant birth weight. In addition, we look at baby's birth length and head circumference which would be related to baby weight. Thus, building a model with predictors of `bhead`, `blength`, `fincome`, `frace`, `mrace`, `ppbmi`, `ppwt`, `smoken`, and `wtgain`, to see the associations of these variables with `bwt`.  

```{r, regression model}
my_model = lm(bwt ~  bhead + blength + fincome +frace + mrace + ppbmi + ppwt + smoken + wtgain, data = birthweight) ## regression model 

my_model %>% ## tidy the regression model
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  mutate(
    term = str_replace(term, "^mrace", "mother race: ")
    ) %>% 
  knitr::kable(digits = 3)
```

Add residuals and predicted values to the model, and plot the model.

```{r, resid and pred}
resid_pred_model = birthweight %>%  ## add residuals and predicted values
  add_residuals(my_model) %>% 
  add_predictions(my_model)

resid_pred_model %>% 
  ggplot(aes(x = pred, y = resid, color = resid)) + 
  geom_point(alpha = .5) + 
  labs(
    title = "Residuals vs fitted values",
    x = "Predicted values",
    y = "Residuals"
  )
  
```

##### Comparison

Now, compare my model to two others. 

```{r, comparison}
## comparisons between models: cross validation
cv_df = 
  crossv_mc(birthweight, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )

cv_df = cv_df %>% 
  mutate(
    my_model  = map(train, ~lm(bwt ~  bhead + blength + fincome +frace + mrace + ppbmi + ppwt + smoken + wtgain, data = .x)),
    model1  = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)), ## length at birth and gestational age as predictors
    model2  = map(train, ~lm(bwt ~ bhead + blength + babysex + bhead * blength + bhead * babysex + blength * bhead + bhead * blength * babysex, data = .x))  ## head circumference, length, sex, and all interactions between these
  ) %>% 
  mutate(
    rmse_my_model = map2_dbl(my_model, test, ~rmse(model = .x, data = .y)),
    rmse_model1 = map2_dbl(model1, test, ~rmse(model = .x, data = .y)),
    rmse_model2 = map2_dbl(model2, test, ~rmse(model = .x, data = .y))
  )

## visualization the comparisons of models
cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse, fill = model)) + 
  geom_violin() + 
  labs(
    title = "Comparisons on RMSE among three models",
    x = "Model",
    y = "RMSE") + 
  scale_x_discrete(
    labels = c("my model", "length & gestational age", "head & length & sex & interaction"))
```

Comparing three models, we found that the model that we created `bwt ~  bhead + blength + fincome +frace + mrace + ppbmi + ppwt + smoken + wtgain` most accurately predicted the baby birthweight. 

